\documentclass{abrice}

\title{Math 352: Midterm 2}
\author{Anthony Brice}

\begin{document}
\maketitle

\section{Exercise 5}

\emph{Suppose we have four chests each having two drawers. Chests $1$
  and $2$ have a gold coin in one drawer and a silver coin in another
  drawer. Chest 3 has two gold coins and Chest 4 has two silver
  coins. A chest is selected at random and a drawer, chosen randomly
  from the chest, is opened. It is found to contain a gold coin. Find
  the probability that the other drawer has
  \begin{enumerate}[label=(\alph*)]
  \item a silver coin;
  \item a gold coin.
  \item What is the probability that Chest $1$ was selected?
  \item What is the probability that Chest $4$ was selected?
  \end{enumerate}
}

\bigskip

\begin{enumerate}[label=(\alph*)]
\item We want to find the probability that the other drawer contains a
  silver coin given that the opened drawer contained a gold
  coin. Since there are $4$ gold coins and $4$ silver coins, each with
  an equal probability of being chosen, the probability of the given
  event is $1/2$. Then we have left to find the probability of the
  intersection of the events that the opened drawer contains a gold
  coin and the other drawer contains a silver coin. Since there are
  eight possible events when selecting a chest then drawer and only two
  of those events involve choosing a chest and drawer that contains a
  gold coin with a silver in the unopened drawer, the probability of
  the intersection is $1/4$. Then the probability in question is
  \[
  { {1 \over 4} \over {1 \over 2}} = {1 \over 2}.
  \]

\item Since the given probability remains the same, we have left to
  find the probability of the intersection of the events that the
  opened drawer contains a gold coin and the other drawer contains a
  gold coin. Since two events of the eight satisfy the conditions, the
  probability of the intersection is $1/4$. Then the probability in
  question is
  \[
  { {1 \over 4} \over {1 \over 2} } = {1 \over 2}.
  \]

\item We want to find the probability that chest $1$ was selected
  given that the chosen drawer contained a gold coin. Our given
  probability remains the same. We want to find the probability of the
  intersection of the events that chest 1 was selected and a gold coin
  was chosen. Since only one out of six events satisfies, the
  probability of the intersection is $1/8$. Then the probability in
  question is $1/4$.

\item Since we are given that a gold coin was found, the probability
  that chest $4$ was selected is certainly $0$.
\end{enumerate}

\section{Exercise 6}

\emph{Suppose $X$ is a random variable with Poisson distribution with
  parameter $\lambda$. Compute the expectation of ${(1 +
  X)}^{-1}$. Hint: What is the Taylor expansion of $e^x$?
}

\bigskip

\begin{align*}
  E\left( {(1+X)}^{-1} \right)
  &= \sum_{x=0}^{\infty} {1 \over 1 + x} \cdot { \lambda^x
    e^{-\lambda} \over x! }\\
  &= e^{-\lambda} \sum_{x=0}^{\infty} {1 \over 1 + x} \cdot {
    \lambda^x \over x!}\\
  &= {1 \over \lambda} e^{-\lambda} \sum_{x=0}^{\infty} { \lambda
    \over 1 + x} \cdot { \lambda^x \over x! }\\
  &= {1 \over \lambda} e^{-\lambda} \sum_{x=0}^{\infty} {
    \lambda^{x+1} \over (x+1)! }\\
  &= {1 \over \lambda} \cdot e^{-\lambda} \cdot e^{\lambda}
    \mathnote{By the Taylor expansion of $e^x$.}\\
  &= {1 \over \lambda}.
\end{align*}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
